Echoes Through Generations: The Permanence of Child Mortality in Moroccan Lineages

### 2024-02-13 outputs/paper/paper.qmd
wrote data section draft

install.packages("gitcreds")

library(tidyverse)
library(palmerpenguins)
library("gitcreds")

@@

Lu and Vogl’s study uncovers a potent and persistent connection between the child mortality experiences of grandmothers and their daughters, drawing on data from 119 Demographic and Health Surveys. In Morocco, the analysis sought to fill the critical gap in understanding the specific magnitude and characteristics of this relationship. The estimands in the study are primarily odds ratios and incidence rate ratios, which provide a measure of the relationship between a mother's experience of sibling death in childhood and her own child's mortality risk. In addition to these, the researchers computed average marginal effects to capture the absolute persistence of child mortality risk. By employing robust statistical techniques and controlling for potential confounders, this paper aims to provide a clear picture of intergenerational child mortality trends within Moroccan society.

@@

# Data {#sec-data}

>>>>

install.packages("ggplot2")

library(ggplot2)

@@

Lu and Vogl’s study uncovers a potent and persistent connection between the child mortality experiences of grandmothers and their daughters, drawing on data from 119 Demographic and Health Surveys. In Morocco, the analysis sought to fill the critical gap in understanding the specific magnitude and characteristics of this relationship. The estimands in the paper are primarily odds ratios and incidence rate ratios, which provide a measure of the relationship between a mother's experience of sibling death in childhood and her own child's mortality risk. In addition to these, the researchers computed average marginal effects to capture the absolute persistence of child mortality risk. By employing robust statistical techniques and controlling for potential confounders, this paper aims to provide a clear picture of intergenerational child mortality trends within Moroccan society.

@@

# Data: Focusing on Morocco's Intergenerational Child Mortality {#sec-data}

The dataset for Lu and Vogl's study provides a detailed snapshot of child mortality rates across generations within Morocco. It encompasses demographic surveys capturing both historical and contemporary data, spanning several decades to allow for the assessment of persistence in child mortality rates from grandmothers to mothers.

The variables within this dataset were carefully selected to reflect key indicators of health outcomes, socioeconomic status, and geographic identifiers for regional analysis within Morocco. These variables include include the number and ages of children who died, causes of death, maternal health indicators, education levels, household income, and access to health services, which are vital for understanding the intergenerational dynamics of child mortality.

Data cleaning was especially important in this case, considering the challenges inherent in longitudinal and generational health data. This ensured the integrity and continuity of data across different time points and handling any missing or inconsistent records, which are typical in historical datasets.

Summary statistics revealed the mean and median child mortality rates, the variance in socio-economic indicators among the surveyed families, and other health-related metrics. This aggregated data provided a foundation for understanding the base conditions of Moroccan families and any preliminary patterns indicating the persistence of child mortality across generations.

Visual representations, such as graphs and tables, illustrated the distribution of mortality rates, the relationship between socio-economic factors and child mortality, and the regional differences within Morocco. These visuals helped to convey the depth and nuances of the intergenerational persistence observed in the Moroccan data.

The appendices include detailed statistical tables, additional graphs showing complex data relationships, and extensive notes on the data cleaning and processing methodology.

This description of the dataset was written with the general ethical practices of demographic and health research in mind and provides a narrative for a subset of data focused on Morocco.

The precise quantification of child mortality rates captured not only the incidence of deaths but also the timing and causes. This entailed using measurement techniques and standard demographic health indicators like the under-five mortality rate (U5MR) or neonatal mortality rate (NMR), which are internationally recognized measures.

Socioeconomic status was gauged using a composite index combining variables such as income, education, and occupation. These measurements were tailored to the Moroccan context to accurately reflect local economic conditions and societal structures.

Demographic characteristics were measured using standard metrics such as age, sex, household size, and variables that capture migration or urban-rural residency, given Morocco's diverse landscape from bustling cities to rural areas.

Measurement in this dataset was meticulous to ensure comparability across generations. For example, the paper needed to account for changes over time in how mortality or socioeconomic status is recorded and reported within Morocco. The potential for measurement error or bias would necessitate rigorous validation techniques, such as cross-referencing with other data sources or conducting field verifications.

The dataset also included constructed variables, which are derived from raw data through transformations or computations to better capture the concepts under study. For instance, a constructed variable might be a mortality risk score that combines the number of sibling deaths with the ages at which they occurred.

In the context of Morocco, the discussion of measurement would also address the cultural and linguistic considerations in the data collection process. Questionnaires and surveys needed to be accurately translated and culturally adapted to ensure the measurement is valid and reliable across different Moroccan communities.

Incorporating a thorough discussion of measurement in relation to the dataset is essential to validate the paper's findings and to ensure that conclusions drawn are based on accurate and reliable data. This discussion would ultimately contribute to the paper's credibility and the robustness of its policy recommendations for Morocco.
